---
title: "Qmm final Project"
author: "Praneeth simha Mallenahalli vijaya,Shiva chaitanya goud gadila,Gayathri yenigalla"
date: "2023-12-10"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#loading the libraries
```{r}
library(dplyr)
library(lpSolve)
```

```{r}
library(dplyr)

# Create the class_students data frame
class_students <- data.frame(
  Student_ID = 1:15,
  Name = c("Student1", "Student2", "Student3", "Student4", "Student5",
           "Student6", "Student7", "Student8", "Student9", "Student10",
           "Student11", "Student12", "Student13", "Student14", "Student15"),
  GPA = c(2.8, 2.4, 3.8,3.7,3,3.2,2,2.2,2.6,3.2,3.6,3.4,2.6,2.9,2.1),  # Random GPAs between 2.0 and 4.0
  Participation_Marks = c(8.1, 6.2, 5.7, 8.8, 5.6, 9, 5.3, 7.7, 7.7, 7.9, 8.1, 8.3, 5.2, 7.8, 7.4),
  Activity = c(10, 5, 10, 5, 10, 6, 7, 7, 6, 5, 9, 10, 8, 7, 5),
  Gender = sample(c("Male", "Female"), 15, replace = TRUE)
)

# Calculate FinalMarks and create class_students_COFF data frame
class_students_COFF <- class_students %>%
  mutate(FinalMarks = ((GPA / 4) * 50) + ((Activity / 10) * 25) + (Participation_Marks * 25))

# View the resulting data frame
View(class_students_COFF)

```

Let:

$x_{ij}$ = 1 if person $j$ belong to group $i$, and 0 if not.  

+ $j$ refers to person = 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15

+ $i$ means group = 1, 2, 3, 4, 5

Objective function: 

\begin{equation*}
  \begin{split}
    Max \: Z &= 198\sum_{i=1}^{5}x_{i1} + 243\sum_{i=1}^{5}x_{i2} + 253\sum_{i=1}^{5}x_{i3} +  248\sum_{i=1}^{5}x_{i4} + 282\sum_{i=1}^{5}x_{i5} + 216\sum_{i=1}^{5}x_{i6} \\
    &+ 250\sum_{i=1}^{5}x_{i7} + 225\sum_{i=1}^{5}x_{i8} + 230\sum_{i=1}^{5}x_{i9} + 231\sum_{i=1}^{5}x_{i10} + 270\sum_{i=1}^{5}x_{i11} +205\sum_{i=1}^{5}x_{i12} + 212\sum_{i=1}^{5}x_{i13} + 183\sum_{i=1}^{5}x_{i14} +207\sum_{i=1}^{5}x_{i15}
    \end{split}
\end{equation*}
 
 Groups

$$ \sum_{j=1}^{15}x_{ij} = 3, where \: i = 1, 2, 3, 4, 5 \: and  \: j = 1,2,3,4, 5, ... , 12, 13, 14, 15$$

 GPA

$$ \sum_{j=1}^{15}GPA_{j}x_{ij}  \geq 15 - i, where \: i = 1, 2, 3, 4,5 \: and  \: j = 1,2,3,4, 5, ... , 12,13,14,15 $$

#Subject to Constraints

#GROUP CONSTRAINT

Group 1: $\sum_{j=1}^{15}G_{i1} = 3$

Group 2: $\sum_{j=1}^{15}G_{i2} = 3$

Group 3: $\sum_{j=1}^{15}G_{i3} = 3$

Group 4: $\sum_{j=1}^{15}G_{i4} = 3$

Group 5: $\sum_{j=1}^{15}G_{i5} = 3$


#STUDENT CONSTRAINTS

Student 1:$\sum_{j=1}^{5}G_{i1} = 1$

Student 2:$\sum_{j=1}^{5}G_{i2} = 1$ 

Student 3:$\sum_{j=1}^{5}G_{i3} = 1$

Student 4:$\sum_{j=1}^{5}G_{i4} = 1$

Student 5:$\sum_{j=1}^{5}G_{i5} = 1$

Student 6:$\sum_{j=1}^{5}G_{i6} = 1$

Student 7:$\sum_{j=1}^{5}G_{i7} = 1$

Student 8:$\sum_{j=1}^{5}G_{i8} = 1$

Student 9:$\sum_{j=1}^{5}G_{i9} = 1$

Student 10:$\sum_{j=1}^{5}G_{i10} = 1$

Student 11:$\sum_{j=1}^{5}G_{i11} = 1$

Student 12:$\sum_{j=1}^{5}G_{i12} = 1$

Student 13:$\sum_{j=1}^{5}G_{i13} = 1$

Student 14:$\sum_{j=1}^{5}G_{i14} = 1$

Student 15:$\sum_{j=1}^{5}G_{i15} = 1$


#GPA CONSTRAINT:

$G_{i,j}$ represents GPA of each student in the group where i represents student and j represents group

Group 1: $2.8G_{11}+2.4G_{21}+3.8G_{31}+3.7G_{41}+3.0G_{51}+3.2G_{61}+2.0G_{71}+2.2G_{81}+2.6G_{91}+3.2G_{101}+3.6G_{ 11}+3.4G_{121}+2.6G_{131}+2.9G_{141}+2.1G_{151}\ge2.9$

Group 2: $2.8G_{12}+2.4G_{22}+3.8G_{32}+3.7G_{42}+3.0G_{52}+3.2G_{62}+2.0G_{72}+2.2G_{82}+2.6G_{92}+3.2G_{102}+3.6G_{112}+3.4G_{122}+2.6G_{132}+2.9G_{142}+2.1G_{152}\ge2.9$

Group 3: $2.8G_{13}+2.4G_{23}+3.8G_{33}+3.7G_{43}+3.0G_{53}+3.2G_{63}+2.0G_{73}+2.2G_{83}+2.6G_{93}+3.2G_{103}+3.6G_{113}+3.4G_{123}+2.6G_{133}+2.9G_{143}+2.1G_{153}\ge2.9$

Group 4: $2.8G_{14}+2.4G_{24}+3.8G_{34}+3.7G_{44}+3.0G_{54}+3.2G_{64}+2.0G_{74}+2.2G_{84}+2.6G_{94}+3.2G_{104}+3.6G_{114}+3.4G_{124}+2.6G_{134}+2.9G_{144}+2.1G_{154}\ge2.9$

Group 5: $2.8G_{15}+2.4G_{25}+3.8G_{35}+3.7G_{45}+3.0G_{55}+3.2G_{65}+2.0G_{75}+2.2G_{85}+2.6G_{95}+3.2G_{105}+3.6G_{115}+3.4G_{125}+2.6G_{135}+2.9G_{145}+2.1G_{155}\ge2.9$


#ACTIVITY CONSTRAINT:

Group 1: $10G_{11}+5G_{21}+10G_{31}+5G_{41}+10G_{51}+6G_{61}+7G_{71}+7G_{81}+6G_{91}+5G_{101}+9G_{111}+10G_{121}+8G_{131}+7G_{141}+5G_{151}\ge7.3$

Group 2: $10G_{12}+5G_{22}+10G_{32}+5G_{42}+10G_{52}+6G_{62}+7G_{72}+7G_{82}+6G_{92}+5G_{102}+9G_{112}+10G_{122}+8G_{132}+7G_{142}+5G_{152}\ge7.3$

Group 3: $10G_{13}+5G_{23}+10G_{33}+5G_{43}+10G_{53}+6G_{63}+7G_{73}+7G_{83}+6G_{93}+5G_{103}+9G_{113}+10G_{123}+8G_{133}+7G_{143}+5G_{153}\ge7.3$

Group 4: $10G_{14}+5G_{24}+10G_{34}+5G_{44}+10G_{54}+6G_{64}+7G_{74}+7G_{84}+6G_{94}+5G_{104}+9G_{114}+10G_{124}+8G_{134}+7G_{144}+5G_{154}\ge7.3$

Group 5: $10G_{15}+5G_{25}+10G_{35}+5G_{45}+10G_{55}+6G_{65}+7G_{75}+7G_{85}+6G_{95}+5G_{105}+9G_{115}+10G_{125}+8G_{135}+7G_{145}+5G_{155}\ge7.3$


#PARTICIPATION CONSTRAINT:

Group 1: $8.1G_{11}+6.2G_{21}+5.7G_{31}+8.8G_{41}+5.6G_{51}+9.0G_{61}+5.3G_{71}+7.7G_{81}+7.7G{91}+7.9G_{101}+8.1G_{111}+8.3G_{121}+5.2G_{131}+7.8G_{141}+7.4G_{151}\ge7.25$

Group 2: $8.1G_{12}+6.2G_{22}+5.7G_{32}+8.8G_{42}+5.6G_{52}+9.0G_{62}+5.3G_{72}+7.7G_{82}+7.7G_{92}+7.9G_{102}+8.1G_{112}+8.3G_{122}+5.2G_{132}+7.8G_{142}+7.4G_{152}\ge7.25$

Group 3: $8.1G_{13}+6.2G_{23}+5.7G_{33}+8.8G_{43}+5.6G_{53}+9.0G_{63}+5.3G_{73}+7.7G_{83}+7.7G_{93}+7.9G_{103}+8.1G_{113}+8.3G_{123}+5.2G_{133}+7.8G_{143}+7.4G_{153}\ge7.25$

Group 4: $8.1G_{14}+6.2G_{24}+5.7G_{34}+8.8G_{44}+5.6G_{54}+9.0G_{64}+5.3G_{74}+7.7G_{84}+7.7G_{94}+7.9G_{104}+8.1G_{114}+8.3G_{124}+5.2G_{134}+7.8G_{144}+7.4G_{154}\ge7.25$

Group 5: $8.1G_{15}+6.2G_{25}+5.7G_{35}+8.8G_{45}+5.6G_{55}+9.0G_{65}+5.3G_{75}+7.7G_{85}+7.7G_{95}+7.9G_{105}+8.1G_{115}+8.3G_{125}+5.2G_{135}+7.8G_{145}+7.4G_{155}\ge7.25$


#Setting the directory 
```{r}
setwd("C:\\Users\\simha\\OneDrive\\Desktop")
```

#Importing the Lp file
```{r}
library(lpSolveAPI)
qmm<- read.lp("qmmfinal.lp")
qmm
```


```{r}
set.type(qmm,1:75,"binary")
```


```{r}
solve(qmm)
```

```{r}
get.objective(qmm)
```

```{r}
get.constraints(qmm)
```

```{r}
m <- get.variables(qmm)
m
```

#optimal distribution of the students
```{r}
x <- matrix(m,nrow=15,byrow=FALSE)
colnames(x) = c("G1","G2","G3","G4","G5")
rownames(x) = class_students_COFF$StudentID
x
```

#GROUPS 	STUDENTS 
#GROUP 1 	Student  1, Student 9 , Student 14 
#GROUP 2	Student 7, Student 8, Student 13
#GROUP 3 	Student 2, Student12 , Student15 
#GROUP 4 	Student 4, Student 6, Student 10 
#GROUP 5	Student 3, Student 5 , Student 11











